---
title: 'Class 8: Machine Learning'
author: "Sandy Schweis"
date: "4/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering

Let's start with an example of running the **k-means()** function

```{r}
# Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))

plot(x)
```


Use the kmeans() function setting k to 2 and nstart=20

```{r}
km <- kmeans(x, centers=2, nstart=20)
```

Inspect/print the results

> Q. How many points are in each cluster?
30

> Q. What ‘component’ of your result object details
 - cluster size? 
 
```{r}
km$size 
```
 
 - cluster assignment/membership?
 
```{r}
km$cluster
```
 
 - cluster center?
 
```{r}
km$centers
```
 
> Plot x colored by the kmeans cluster assignment and
 add cluster centers as blue points
 
```{r}
km <- kmeans(x, centers=2, nstart=20)
```
```{r}
km
```
 
```{r}
plot(x, col= km$cluster)
points(km$centers, pch=18, col="purple", cex=3)
```
 
 ## Hierarchical clustering example
 
 We must give the **hclust()** function a distance matrix not the raw data as input
 
```{r}
# distance matrix calculation
d <- dist(x)

# clustering
hc <- hclust(d)
plot(hc)
```
 
```{r}
cutree(hc, h=6)
```
 
```{r}
cutree(hc, k=2)
```
 
```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")

# Step 2. Plot the data without clustering
plot(x)

# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )

plot(x, col=col)
```
 
 
>Q. Use the dist(), hclust(), plot() and cutree()
 functions to return 2 and 3 clusters
>Q. How does this compare to your known 'col' groups?

```{r}
hc <- hclust(dist(x))
gp2 <- cutree(hc, k=2)
plot(hc)
abline(h=2, col= "pink")
abline(h=2.8, col= "purple")
```
 
```{r}
table(gp2)
```


# PCA: Principal Component Analysis

We will use the **prcomp()** function for PCA
```{r}
## You can also download this file from the class website!
mydata <- read.csv("https://tinyurl.com/expression-CSV",
 row.names=1)
head(mydata, 10) 
```

Running PCA analysis on the transpose of our data
```{r}
pca <- prcomp(t(mydata), scale= TRUE)
```

PCA plot 
```{r}
plot(pca$x[,1], pca$x[,2], xlab = "PC1", ylab = "PC2") 
```
 
 
```{r}
## Precent variance is often more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1) 

head(pca.var.per)
```
 
```{r}
barplot(pca.var.per, main="Scree Plot",
 xlab="Principal Component", ylab="Percent Variation")
```
 
```{r}
## A vector of colors for wt and ko samples
colvec <- colnames(mydata)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
 xlab=paste0("PC1 (", pca.var.per[1], "%)"),
 ylab=paste0("PC2 (", pca.var.per[2], "%)"))
```
```{r}
x <- read.csv("UK_foods.csv")
```
 
```{r}
col(x)
```


```{r}
dim(x)
```

```{r}
rownames(x) <- x[,1]
x <- x[,-1]
head(x)
```


```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```
 
```{r}
# Use the prcomp() PCA function 
pca <- prcomp( t(x) )
summary(pca)
```
 
```{r}
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500))
text(pca$x[,1], pca$x[,2], colnames(x), col = rainbow(ncol(x)))
```
 
```{r}
v <- round( pca$sdev^2/sum(pca$sdev^2) * 100 )
v
```

```{r}
## or the second row here...
z <- summary(pca)
z$importance
```

```{r}
barplot(v, xlab="Principal Component", ylab="Percent Variation")
```

```{r}
## Lets focus on PC1 as it accounts for > 90% of variance 
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```


 
 
 
 
 
 