---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE, row.names = 1)
metadata <- read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```

```{r}
head(counts)
```

```{r}
head(metadata)
```

Let's check the correspondence of our countData cols and our metaData rows. Will be required for DESeq later.

```{r}
all(colnames(counts) == metadata$id)
```


```{r}
control <- metadata[metadata[,"dex"]=="control",]
control.mean <- rowSums( counts[ ,control$id] )/length(control$id)
names(control.mean) <- counts$ensgene
```

Determine the mean count values for all genes accross control experiments
```{r}
control.mean <- rowSums( counts[ ,control$id] )/length(control$id)
```

```{r}
treated.inds <- metadata$dex == "treated"
treated <- metadata[treated.inds, ]
treated
```


```{r}
treated.mean <- rowSums(counts[, treated$id])/ length(treated$id)
```


Let's store the control.mean and treated.mean together
```{r}
meancounts <- data.frame(control.mean, treated.mean)
```


What is the sum counts across all genes in control and treated?
```{r}
colSums(meancounts)
```

```{r}
plot(meancounts)
```

Plot on a log-log scale so we can see our data
```{r}
plot(meancounts, log= "xy")
```

Calculate log2foldchange, add it to our meancounts data.frame and inspect the results either with the head() or view() function for example. 
```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```


Filter out and remove NaN and -Ing values. arr.ind treats data as one big matrix.
```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)
to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```

```{r}
nrow(mycounts)
```

How many genes are up regulated in the drug treated cells and how many are down regulated?
```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```

```{r}
sum(up.ind)
sum(down.ind)
```

```{r}
anno <- read.csv("annotables_grch38.csv")
head(anno)
```

Use **merge()** function to add the annotation data from the 'anno' object to our RNA-seq results in 'mycounts'
```{r}
# use the merge function
mycounts.anno <- merge(mycounts, anno, by.x = "row.names" , by.y = "ensgene")
mycounts.anno
```

```{r}
library("AnnotationDbi")
```

```{r}
library("org.Hs.eg.db")
```

```{r}
columns(org.Hs.eg.db)
```

Use **mapIds()** function from bioconductor to add annotation data to our results
```{r}
mycounts$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
```

```{r}
head(mycounts)
```

```{r}
mycounts$entrez <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="ENTREZID",
                     keytype="ENSEMBL")
```


```{r}
head(mycounts)
```


##Use DESeq2 
```{r}
library(DESeq2)
citation("DESeq2")
```

```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <- read.csv("airway_metadata.csv", stringsAsFactors = FALSE)
```

```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
head(res)
```

```{r}
summary(res)
```

```{r}
res05 <- results(dds, alpha=0.05)
summary(res05)
```

See our results as a data.frame object
```{r}
res01 <- as.data.frame(results(dds, alpha=0.01))
head(res01)
```


Let's sort or order our results by the adjusted p-value
```{r}
ord.inds <- order(res01$padj)
write.csv(res01[ord.inds,], file="signif01_results.csv")
```


## Volcano plot
(results are contained in res01)
```{r}
mycols <- rep("gray", nrow(res01))
mycols[ res01$padj <0.01] <- "black"
mycols[ abs(res01$log2FoldChange) > 2 ] <- "pink"
mycols[ (res01$padj <0.01) & abs(res01$log2FoldChange) > 2] <- "purple"

plot(res01$log2FoldChange, -log(res01$padj), xlab="Fold Change (Drug/Control)", ylab= "-log P-value", col=mycols)
```











