---
title: 'Class 7: R Functions and Packages'
author: "Sandy Schweis"
date: "4/23/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Functions Revisted 

We will source a file from online with our functions from last day 
```{r}
source("http://tinyurl.com/rescale-R")
```

Try out the last day's rescale() function

```{r}
rescale(1:10)
```

Try the rescale2() function that catches string inputs

#Find missing NA values into vectors

Start with a simple example of the larger problem I am trying to solve
```{r}
x <- c(1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)
```

```{r}
is.na(x)
```

```{r}
is.na(y)
```

Try putting these together with an AND
```{r}
is.na(x) & is.na(y)
```

Take the sum() to find out how many TRUE values we have and thus how many NAs we have in both x and y 

```{r}
sum(is.na(x) & is.na(y))
```

Now I can this into a function...

```{r}
both_na <- function(x, y) {
  sum(is.na(x) & is.na(y))
}
```


```{r}
both_na(x, y)
```

```{r}
both_na(x, c(NA,3, NA, 2, NA))
```

Try both_na2() function on our different length input vectors

```{r}
both_na2(x, y)
```
##Intersect function

```{r}
df1
```

```{r}
df2
```

Make things simple
```{r}
x <- df1$IDs
y <- df2$IDs

x
y
```

```{r}
intersect(x, y)
```

```{r}
which(x %in% y)
```


```{r}
x[x %in% y]
y[y %in% x]
```


```{r}
cbind(x[x %in% y],
      y[y %in% x])
```

Use RStudio shortcut CODE > EXTRACT FUNCTION to turn our snippet into a working function

```{r}
gene_intersect <- function(x, y) {
  cbind(x[x %in% y],
        y[y %in% x])
}
```

```{r}
gene_intersect2(df1, df2)
```


```{r}
gene_intersect3(df1, df2)
```

```{r}
merge(df1, df2, by = "IDs")
```

Find average grade dropping the worst homework score
```{r}
a <- c(100, 100, 100, 100, 100, 100, 100, 90)
b <- c(100, 90, 90, 90, 90, 90, 97, 80)
```


```{r}
grade <- function(a) {
  (sum(a) - min(a)) / (length(a)-1)
}
```

```{r}
grade(a)
```

```{r}

```





















